---
title: "TMSstroke_SICI"
output: html_document
date: "2023-04-11"
---

## Load required R packages 
```{r setup, include=FALSE}
library(knitr)
opts_chunk$set(echo = TRUE)

library(readr)
library(tidyverse)
library(ggpubr)
library(rstatix)
library(dplyr)
```

## General
State 1: No SICI (just test pulse)
State 2: SICI (conditioning and test pulse)

Baycrest data:

subID Affected Hemisphere
18989	Left
13108	Right
19685	Left
19672	Left
19673	Right
19674	Right
19746	Right
19684	Right
19747	Right
n = 9

McMaster data:

## Load data
```{r}
setwd("/Users/connectome/Documents/Postdoc/Baycrest/TMS_stroke/repositories")

# data is already in long format so don't need to do anything to it
rawdataSICI <- read_csv("data/SICI_processed_R.csv") #read in the raw data files
head(rawdataSICI, 9) #print the first 9 rows

rawdataSICI.amp <- rawdataSICI %>% 
  select(1,2,5)
head(rawdataSICI.amp, 10)

dataSICI.amp <- pivot_wider(rawdataSICI.amp, subID, names_from = hemisphere_treatment_state, values_from = MEP_amp) %>%
  transmute(
    subID, 
    affected_SICI_pre = affected_pre_2 / affected_pre_1,
    affected_SICI_post = affected_post_2 / affected_post_1, 
    unaffected_SICI_pre = unaffected_pre_2 / unaffected_pre_1, 
    unaffected_SICI_post = unaffected_post_2 / unaffected_post_1
  ) %>%
  pivot_longer(-subID, names_to = "hemisphere_treatment_state", values_to = "MEP_amp")

# export to csv
write.csv(dataSICI.amp, "/Users/connectome/Documents/Postdoc/Baycrest/TMS_stroke/repositories/data/SICI_amplitude_rmcorr.csv",
          row.names=FALSE)

rawdataSICI.latency <- rawdataSICI %>% 
  select(1,2,11)
head(rawdataSICI.latency, 10)

dataSICI.latency <- pivot_wider(rawdataSICI.latency, subID, names_from = hemisphere_treatment_state, values_from = MEP_latency) %>%
  transmute(
    subID, 
    affected_SICI_pre = affected_pre_2 / affected_pre_1,
    affected_SICI_post = affected_post_2 / affected_post_1, 
    unaffected_SICI_pre = unaffected_pre_2 / unaffected_pre_1, 
    unaffected_SICI_post = unaffected_post_2 / unaffected_post_1
  ) %>%
  pivot_longer(-subID, names_to = "hemisphere_treatment_state", values_to = "MEP_latency")

#export to csv
write.csv(dataSICI.latency, "/Users/connectome/Documents/Postdoc/Baycrest/TMS_stroke/repositories/data/SICI_latency_rmcorr.csv",
          row.names=FALSE)
```

## Summary statistics and pre-test checks
```{r}
rawdataSICI %>%
  group_by(hemisphere_treatment_state) %>%
  get_summary_stats(MEP_amp, type = "mean_sd")

rawdataSICI %>%
  group_by(hemisphere_treatment_state) %>%
  get_summary_stats(MEP_latency, type = "mean_sd")

# Get data in wide format for normality testing below
dataSICI.amp.wide <- rawdataSICI.amp %>% 
    pivot_wider(names_from = hemisphere_treatment_state, values_from = MEP_amp)
head(dataSICI.amp.wide, 10)

dataSICI.latency.wide <- rawdataSICI.latency %>% 
    pivot_wider(names_from = hemisphere_treatment_state, values_from = MEP_latency)
head(dataSICI.latency.wide, 10)

# Test for normality, if p > 0.05 can assume normality
shapiro.test(as.numeric(as.character(dataSICI.amp.wide$affected_pre_1))) # => p = 0.9435
shapiro.test(as.numeric(as.character(dataSICI.amp.wide$affected_pre_2))) # => p = 0.4708
shapiro.test(as.numeric(as.character(dataSICI.amp.wide$unaffected_pre_1))) # => p = 0.4174
shapiro.test(as.numeric(as.character(dataSICI.amp.wide$unaffected_pre_2))) # => p = 0.7454
shapiro.test(as.numeric(as.character(dataSICI.amp.wide$affected_post_1))) # => p = 0.8725
shapiro.test(as.numeric(as.character(dataSICI.amp.wide$affected_post_2))) # => p = 0.6755
shapiro.test(as.numeric(as.character(dataSICI.amp.wide$unaffected_post_1))) # => p = 0.1065
shapiro.test(as.numeric(as.character(dataSICI.amp.wide$unaffected_post_2))) # => p = 0.1768

shapiro.test(as.numeric(as.character(dataSICI.latency.wide$affected_pre_1))) # => p = 0.3812
shapiro.test(as.numeric(as.character(dataSICI.latency.wide$affected_pre_2))) # => p = 0.6911
shapiro.test(as.numeric(as.character(dataSICI.latency.wide$unaffected_pre_1))) # => p = 0.2679
shapiro.test(as.numeric(as.character(dataSICI.latency.wide$unaffected_pre_2))) # => p = 0.08054
shapiro.test(as.numeric(as.character(dataSICI.latency.wide$affected_post_1))) # => p = 0.8355
shapiro.test(as.numeric(as.character(dataSICI.latency.wide$affected_post_2))) # => p = 0.8771
shapiro.test(as.numeric(as.character(dataSICI.latency.wide$unaffected_post_1))) # => p = 0.3325
shapiro.test(as.numeric(as.character(dataSICI.latency.wide$unaffected_post_2))) # => p = 0.05646
```

# AMPLITUDE

## Affected hemi amplitude - paired t-test
```{r}
dataSICI.amp.affected <- filter(dataSICI.amp, grepl("^affected", hemisphere_treatment_state))

res.affected.amp <- t.test(MEP_amp ~ hemisphere_treatment_state, 
                           data = na.omit(dataSICI.amp.affected), paired = TRUE)
res.affected.amp

stat.affected.amp <- na.omit(dataSICI.amp.affected)  %>% 
  t_test(MEP_amp ~ hemisphere_treatment_state, paired = TRUE, detailed = TRUE) %>%
  add_significance()
stat.affected.amp

dataSICI.amp.affected %>% cohens_d(MEP_amp ~ hemisphere_treatment_state, paired = TRUE)

# Create a box plot
bxp <- ggpaired(dataSICI.amp.affected, x = "hemisphere_treatment_state", y = "MEP_amp", id = "subID",
                label = "subID", 
         order = c("affected_SICI_pre", "affected_SICI_post"),
         ylab = "Affected MEP Amplitude", xlab = "Pre/Post Treatment")

# Add p-value and significance levels
stat.affected.amp <- stat.affected.amp %>% add_xy_position(x = "hemisphere_treatment_state")
bxp.SICIamp.affected <- bxp + 
  stat_pvalue_manual(stat.affected.amp, tip.length = 0) +
  labs(subtitle = get_test_label(stat.affected.amp, detailed= TRUE))
bxp.SICIamp.affected

ggsave("bxp_SICIamp_affected.tiff", 
       plot = bxp.SICIamp.affected,
       width = 7,
       height = 5
)
```

## Unaffected hemi amplitude - paired t-test
```{r}
dataSICI.amp.unaffected <- filter(dataSICI.amp, grepl("^unaffected", hemisphere_treatment_state))

res.unaffected.amp <- t.test(MEP_amp ~ hemisphere_treatment_state, 
                           data = na.omit(dataSICI.amp.unaffected), paired = TRUE)
res.unaffected.amp

stat.unaffected.amp <- na.omit(dataSICI.amp.unaffected)  %>% 
  t_test(MEP_amp ~ hemisphere_treatment_state, paired = TRUE, detailed = TRUE) %>%
  add_significance()
stat.unaffected.amp

dataSICI.amp.unaffected %>% cohens_d(MEP_amp ~ hemisphere_treatment_state, paired = TRUE)

# Create a box plot
bxp <- ggpaired(dataSICI.amp.unaffected, x = "hemisphere_treatment_state", y = "MEP_amp", id = "subID",
                label = "subID", 
         order = c("unaffected_SICI_pre", "unaffected_SICI_post"),
         ylab = "Unaffected MEP Amplitude", xlab = "Pre/Post Treatment")

# Add p-value and significance levels
stat.unaffected.amp <- stat.unaffected.amp %>% add_xy_position(x = "hemisphere_treatment_state")
bxp.SICIamp.unaffected <- bxp + 
  stat_pvalue_manual(stat.unaffected.amp, tip.length = 0) +
  labs(subtitle = get_test_label(stat.unaffected.amp, detailed= TRUE))
bxp.SICIamp.unaffected

ggsave("bxp_SICIamp_unaffected.tiff", 
       plot = bxp.SICIamp.unaffected,
       width = 7,
       height = 5
)
```

# LATENCY

## Affected hemi latency - paired t-test
```{r}
dataSICI.latency.affected <- filter(dataSICI.latency, grepl("^affected", hemisphere_treatment_state))

res.affected.latency <- t.test(MEP_latency ~ hemisphere_treatment_state, 
                           data = na.omit(dataSICI.latency.affected), paired = TRUE)
res.affected.latency

stat.affected.latency <- na.omit(dataSICI.latency.affected)  %>% 
  t_test(MEP_latency ~ hemisphere_treatment_state, paired = TRUE, detailed = TRUE) %>%
  add_significance()
stat.affected.latency

dataSICI.latency.affected %>% cohens_d(MEP_latency ~ hemisphere_treatment_state, paired = TRUE)

# Create a box plot
bxp <- ggpaired(dataSICI.latency.affected, x = "hemisphere_treatment_state", y = "MEP_latency", id = "subID",
                label = "subID",
         order = c("affected_SICI_pre", "affected_SICI_post"),
         ylab = "Affected MEP Latency", xlab = "Pre/Post Treatment", repel = TRUE)

# Add p-value and significance levels
stat.affected.latency <- stat.affected.latency %>% add_xy_position(x = "hemisphere_treatment_state")
bxp.SICIlatency.affected <- bxp + 
  stat_pvalue_manual(stat.affected.latency, tip.length = 0) +
  labs(subtitle = get_test_label(stat.affected.latency, detailed= TRUE))
bxp.SICIlatency.affected

ggsave("bxp_SICIlatency_affected.tiff", 
       plot = bxp.SICIlatency.affected,
       width = 7,
       height = 5
)
```

## Unaffected hemi latency - paired t-test
```{r}
dataSICI.latency.unaffected <- filter(dataSICI.latency, grepl("^unaffected", hemisphere_treatment_state))

res.unaffected.latency <- t.test(MEP_latency ~ hemisphere_treatment_state, 
                           data = na.omit(dataSICI.latency.unaffected), paired = TRUE)
res.unaffected.latency

stat.unaffected.latency <- na.omit(dataSICI.latency.unaffected)  %>% 
  t_test(MEP_latency ~ hemisphere_treatment_state, paired = TRUE, detailed = TRUE) %>%
  add_significance()
stat.unaffected.latency

dataSICI.latency.unaffected %>% cohens_d(MEP_latency ~ hemisphere_treatment_state, paired = TRUE)

# Create a box plot
bxp <- ggpaired(dataSICI.latency.unaffected, x = "hemisphere_treatment_state", y = "MEP_latency", id = "subID",
                label = "subID",
         order = c("unaffected_SICI_pre", "unaffected_SICI_post"),
         ylab = "Unaffected MEP Latency", xlab = "Pre/Post Treatment", repel = TRUE)

# Add p-value and significance levels
stat.unaffected.latency <- stat.unaffected.latency %>% add_xy_position(x = "hemisphere_treatment_state")
bxp.SICIlatency.unaffected <- bxp + 
  stat_pvalue_manual(stat.unaffected.latency, tip.length = 0) +
  labs(subtitle = get_test_label(stat.unaffected.latency, detailed= TRUE))
bxp.SICIlatency.unaffected

ggsave("bxp_SICIlatency_unaffected.tiff", 
       plot = bxp.SICIlatency.unaffected,
       width = 7,
       height = 5
)
```