---
title: "TMSstroke_RC"
output: html_document
date: "2023-08-04"
---

## Load the required libraries
```{r}
library(readr)
library(dplyr, warn=FALSE)
library(tidyr)
library(ggplot2) 
library(cowplot) 
library(RColorBrewer) 
library(rmcorr)
library(ggrepel)
library(minpack.lm)
```

## Get sigmoid function parameters (max plateau, min plateau, S50, slope) for t-tests
From (DOI: 10.1109/TNSRE.2019.2914475) and (DOI: 10.1007/s00221-004-2134-4), the recruitment curve model for parameter estimation
that captures the parameters of most interest to scientists and clinicians, such as the recruitment curve midpoint, slope, and high
saturation level (plateau). This is defined as:

y = min + [(max - min)/(1 + exp((S50 - x)/slope))]

y = mep amplitude
x = pMSO (TMS stimulator output)
min = minimum amplitude value (min plateau)
max = maximum amplitude value (max plateau)
slope = slope
S50 = midpoint

```{r}
# Get a list of the files to read
files <- list.files("/Users/connectome/Documents/Postdoc/Baycrest/TMS_stroke/processed_data/rc/average/unaffected_hemi", 
                    pattern = "\\.csv", full.names = TRUE)

# Read files and add an id
dfRC_all <- read_csv(files, id = "file") |>
  # Clean file name
  mutate(file = gsub("\\-mepRC-UH-avg.csv", "", basename(file))) |>
  # separate into subject and treatment columns
  separate(file, into = c("subID", "treatment"), sep = "_") |> 
  # Select and rename
  select(subID, treatment, pMSO, mepAMP = "MEP AMP (mV)") |> 
  pivot_wider(names_from = treatment, values_from = mepAMP, names_prefix = "mepAMP_")
dfRC_all

# pivot df into long format
dfRC_long <- dfRC_all %>%
  pivot_longer(contains('mep'), names_sep = '_', 
               names_to = c('.value', 'prepost')) %>%
  mutate(subID = factor(subID), prepost = factor(prepost, c('pre', 'post')))

# plot pre vs post
p <- ggplot(dfRC_long, aes(x = pMSO, y = mepAMP, colour = factor(subID))) +
  geom_point() +
  geom_smooth(method = nlsLM, se = FALSE, 
              formula = y ~ plateau.min + ((plateau.max - plateau.min) / (1 + exp((S50 - x)/slope))),
#              upper = c(plateau = 13, S50 = Inf, slope = Inf),
              method.args = list(start = list(plateau.min = 0, plateau.max = 15, S50 = 65, slope = 8))
             ) +
  facet_grid(.~prepost)
p

# calculate model parameters (plateau max and min, S50, slope) for pre and post amplitude
params <- dfRC_long %>%
  group_by(subID, prepost) %>%
  group_map(.f = ~ nlsLM(mepAMP ~ plateau.min + ((plateau.max - plateau.min) / (1 + exp((S50 - pMSO)/slope))), 
                         data = .x, 
                         start = list(plateau.min = 0, plateau.max = 15, S50 = 65, slope = 8) #,
#                         upper = c(plateau.min = 0, plateau.max = 15, S50 = Inf, slope = Inf)
                         ) %>%
              coef() %>%
              t() %>%
              as.data.frame() %>%
              mutate(pre_or_post = .y$prepost, .before = 1) %>%
              mutate(subID = .y$subID, .before = 2)) %>%
  bind_rows() %>%
  arrange(pre_or_post, subID)
params
```
