---
title: "TMSstroke_RMT"
author: "Alica Rogojin"
output: html_TMS-stroke_RMT
date: "2023-02-14"
---

## Load required R packages 
```{r setup, include=FALSE}
library(knitr)
opts_chunk$set(echo = TRUE)

library(readr)
library(tidyverse)
library(ggpubr)
library(rstatix)
library(dplyr)
```

## General
Baycrest data:

subID Affected Hemisphere
18989	Left
13108	Right
19685	Left
19672	Left
19673	Right
19674	Right
19746	Right
19684	Right
19747	Right
n = 9

McMaster data:

## Load data
```{r}
setwd("/Users/connectome/Documents/Postdoc/Baycrest/TMS_stroke/repositories")

# Wide format
dataRMT <- read_csv("data/TMSstroke_RMT.csv") #read in the raw data files

dataRMT$sex <- factor(dataRMT$sex,
                          level=c(0,1),
                          labels=c("male", "female")
                     ) 

head(dataRMT, 9) #print the first 9 rows

# Transform into long data: gather the before and after values in the same column
dataRMT.affected70.long <- dataRMT %>%
  gather(key = "affected70_time", value = "affected_RMT70", affected_70_pre, affected_70_post)
dataRMT.affected70.long <- select(dataRMT.affected70.long, -5:-11)
head(dataRMT.affected70.long, 10)

dataRMT.unaffected70.long <- dataRMT %>%
  gather(key = "unaffected70_time", value = "unaffected_RMT70", unaffected_70_pre, unaffected_70_post)
dataRMT.unaffected70.long <- select(dataRMT.unaffected70.long, -5:-11)
head(dataRMT.unaffected70.long, 10)

dataRMT.affected50.long <- dataRMT %>%
  gather(key = "affected50_time", value = "affected_RMT50", affected_50_pre, affected_50_post)
dataRMT.affected50.long <- select(dataRMT.affected50.long, -5:-11)
head(dataRMT.affected50.long, 10)

dataRMT.unaffected50.long <- dataRMT %>%
  gather(key = "unaffected50_time", value = "unaffected_RMT50", unaffected_50_pre, unaffected_50_post)
dataRMT.unaffected50.long <- select(dataRMT.unaffected50.long, -5:-11)
head(dataRMT.unaffected50.long, 10)
```

## Summary statistics and pre-test checks
```{r}
dataRMT.affected70.long %>%
  group_by(affected70_time) %>%
  get_summary_stats(affected_RMT70, type = "mean_sd")

dataRMT.unaffected70.long %>%
  group_by(unaffected70_time) %>%
  get_summary_stats(unaffected_RMT70, type = "mean_sd")

dataRMT.affected50.long %>%
  group_by(affected50_time) %>%
  get_summary_stats(affected_RMT50, type = "mean_sd")

dataRMT.unaffected50.long %>%
  group_by(unaffected50_time) %>%
  get_summary_stats(unaffected_RMT50, type = "mean_sd")

# Test for normality, if p > 0.05 can assume normality
shapiro.test(dataRMT$affected_70_pre) # => p = 0.253
shapiro.test(dataRMT$affected_70_post) # => p = 0.9666
shapiro.test(dataRMT$unaffected_70_pre) # => p = 0.3487
shapiro.test(dataRMT$unaffected_70_post) # => p = 0.9576
shapiro.test(dataRMT$affected_50_pre) # => p = 0.1287
shapiro.test(dataRMT$affected_50_post) # => p = 0.821
shapiro.test(dataRMT$unaffected_50_pre) # => p = 0.5966
shapiro.test(dataRMT$unaffected_50_post) # => p = 0.05421
```

## Affected hemi with 70 mm coil - paired t-test
```{r}
res.affected70 <- t.test(affected_RMT70 ~ affected70_time, data = na.omit(dataRMT.affected70.long), paired = TRUE)
res.affected70

stat.affected70 <- na.omit(dataRMT.affected70.long)  %>% 
  t_test(affected_RMT70 ~ affected70_time, paired = TRUE, detailed = TRUE) %>%
  add_significance()
stat.affected70

dataRMT.affected70.long %>% cohens_d(affected_RMT70 ~ affected70_time, paired = TRUE)

# Create a box plot
bxp <- ggpaired(dataRMT.affected70.long, x = "affected70_time", y = "affected_RMT70", id = "subID",
         order = c("affected_70_pre", "affected_70_post"),
         ylab = "RMT", xlab = "Pre/Post Treatment")

# Add p-value and significance levels
stat.affected70 <- stat.affected70 %>% add_xy_position(x = "affected70_time")
bxp + 
  stat_pvalue_manual(stat.affected70, tip.length = 0) +
  labs(subtitle = get_test_label(stat.affected70, detailed= TRUE))

```

## Unaffected hemi with 70 mm coil - paired t-test
```{r}
res.unaffected70 <- t.test(unaffected_RMT70 ~ unaffected70_time, data = na.omit(dataRMT.unaffected70.long), paired = TRUE)
res.unaffected70

stat.unaffected70 <- na.omit(dataRMT.unaffected70.long)  %>% 
  t_test(unaffected_RMT70 ~ unaffected70_time, paired = TRUE, detailed = TRUE) %>%
  add_significance()
stat.unaffected70

dataRMT.unaffected70.long %>% cohens_d(unaffected_RMT70 ~ unaffected70_time, paired = TRUE)

# Create a box plot
bxp <- ggpaired(dataRMT.unaffected70.long, x = "unaffected70_time", y = "unaffected_RMT70", id = "subID",
         order = c("unaffected_70_pre", "unaffected_70_post"),
         ylab = "RMT", xlab = "Pre/Post Treatment")

# Add p-value and significance levels
stat.unaffected70 <- stat.unaffected70 %>% add_xy_position(x = "unaffected70_time")
pRMT_unaffected <- bxp + 
  stat_pvalue_manual(stat.unaffected70, tip.length = 0) +
  labs(subtitle = get_test_label(stat.unaffected70, detailed= TRUE))
stat.unaffected70

ggsave("p_RMT_unaffected.tiff", 
       plot = pRMT_unaffected,
       width = 8,
       height = 7
)
```

## Affected hemi with 50 mm coil - paired t-test
```{r}
res.affected50 <- t.test(affected_RMT50 ~ affected50_time, data = na.omit(dataRMT.affected50.long), paired = TRUE)
res.affected50

stat.affected50 <- na.omit(dataRMT.affected50.long)  %>% 
  t_test(affected_RMT50 ~ affected50_time, paired = TRUE, detailed = TRUE) %>%
  add_significance()
stat.affected50

dataRMT.affected50.long %>% cohens_d(affected_RMT50 ~ affected50_time, paired = TRUE)

# Create a box plot
bxp <- ggpaired(dataRMT.affected50.long, x = "affected50_time", y = "affected_RMT50", id = "subID",
         order = c("affected_50_pre", "affected_50_post"),
         ylab = "RMT", xlab = "Pre/Post Treatment")

# Add p-value and significance levels
stat.affected50 <- stat.affected50 %>% add_xy_position(x = "affected50_time")
bxp + 
  stat_pvalue_manual(stat.affected50, tip.length = 0) +
  labs(subtitle = get_test_label(stat.affected50, detailed= TRUE))

```

## Unaffected hemi with 50 mm coil - paired t-test
```{r}
res.unaffected50 <- t.test(unaffected_RMT50 ~ unaffected50_time, data = na.omit(dataRMT.unaffected50.long), paired = TRUE)
res.unaffected50

stat.unaffected50 <- na.omit(dataRMT.unaffected50.long)  %>% 
  t_test(unaffected_RMT50 ~ unaffected50_time, paired = TRUE, detailed = TRUE) %>%
  add_significance()
stat.unaffected50

dataRMT.unaffected50.long %>% cohens_d(unaffected_RMT50 ~ unaffected50_time, paired = TRUE)

# Create a box plot
bxp <- ggpaired(dataRMT.unaffected50.long, x = "unaffected50_time", y = "unaffected_RMT50", id = "subID",
         order = c("unaffected_50_pre", "unaffected_50_post"),
         ylab = "RMT", xlab = "Pre/Post Treatment")

# Add p-value and significance levels
stat.unaffected50 <- stat.unaffected50 %>% add_xy_position(x = "unaffected50_time")
bxp + 
  stat_pvalue_manual(stat.unaffected50, tip.length = 0) +
  labs(subtitle = get_test_label(stat.unaffected50, detailed= TRUE))

```

## Adjust t-test p-values for multiple comparisons
```{r}
# If only using 70 mm coil results, only need to adjust the p-values for those two t-tests
print(p.adjust(c(0.022, 0.2863), method = "holm")) # => 0.0324, 0.2863
```



